
---
title: "Basic of SpotifyViz"
author: " "
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth : 2
vignette: >
  %\VignetteIndexEntry{Basic of SpotifyViz}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE ,
  warning = FALSE, 
  fig.width = 10, 
  fig.height = 7
  
)
```

```{r, warning=F}
library(spotifyviz)
```

This vignette introduces to basic functionalities of spotifyviz package. 

# Loading data


To analyze data from spotify first we have to load the data.  We're going to use `complete_streaming_history` to load streaming history, `prepare_playlists` to load playlists and `make_search_queries_dt` for search queries. These functions take as an argument the path to the folder where the data from spotify is kept. They assume that all data in this folder is from one spotify user, so if you have data from more than one user keep them in separate folders.

```{r, eval = FALSE}

streaming_history <- complete_streaming_history("path")

search_queries <- make_search_queries_dt("path")

playlists <- prepare_playlists("path")

```


In further examples data provided by the package will be used.

```{r}
streaming_history <- spotifyviz::clean_stream_his
  
search_queries <- spotifyviz::clean_search_queries
  
playlists <- spotifyviz::clean_playlist


```
  

The loaded data looks like this:

```{r}

head(streaming_history)

head(search_queries)

head(playlists)
```

# Preparing the additional data

Now, since we have our data in appropriate form, we can start analyzing them.

### Example

```{r}
# the wide form
# playlist_dt = playlists : a data table containing user playlists, made with prepare_playlis
# streaming_history : = streaming_history : part of streaming history of one user, made with complete_streaming_history

playlists_wide <- str_his_with_playlists_wide(playlists, streaming_history)

# see the data

head(playlists_wide)

# the long form

# playlist_dt = playlists : a data table containing user playlists, made with prepare_playlis
# streaming_history : = streaming_history : part of streaming history of one user, made with complete_streaming_history


playlist_long <- str_his_with_playlists_long(playlists, streaming_history)

# see the data

head(playlist_long)
```


## continuous_listening 

This function adds the "listening number" to the streaming history data table - the number which determines to which listening session (the session - the tracks were played with less than `mins` minutes of breaks between them) the track belongs.

###Example

```{r}
# streaming_history = streaming_history : streaming_history : part of streaming history of one user, made with complete_streaming_history
# mins = 5 : the breaks must be less than 5 minutes
sessions <- continuous_listening(streaming_history, mins = 5)

# see the data
head(sessions)

```


# Stats and summaries of the data

## make_summary_dt

This function makes the table with brief summary of the data filtered to the period of time provided. It uses `measure_how_long_listened` and `count_skipped` functions.

# Example 
```{r}
# streaming_history = streaming_history : part of streaming history of one user, made with complete_streaming_history
# start_date = "2018-01-01" : indicates start of the period of time should be on "2018-01-01"
# end_date = "2019-01-01" : indicates end of the period of time should be on "2019-01-01"
# as_percentage  = TRUE  : tracks skipped and time listened to will be a percentegae of all tracks and whole time period.
make_summary_dt(streaming_history, start_date = "2018-01-01", end_date = "2019-01-01", as_percentage = TRUE)
```

## most_played

This function finds the most played tracks or artist and returns a data table with either track or artist name and the times it was played.

### Example

```{r}
# streaming_history = streaming_history : part of streaming history of one user, made with complete_streaming_history
# track_or_artist = "artist" : count how many times each artist was played
# show_skipped = FALSE : don't include skipped tracks in the count
# how_many = 10 : show first ten results

most_played(streaming_history, track_or_artist = "artist", show_skipped = FALSE, how_many = 10)
```


## most_skipped

This function finds the most skipped  (their play time was less than 10s) tracks or artist and returns a data table with either artist or track names and the times they were skipped.

### Example

```{r}
# streaming_history = streaming_history : part of streaming history of one user, made with complete_streaming_history
# track_or_artist = "artist" : count how many times each artist was played
# for_view = TRUE : show results in format good for reading but not the best for futher computations 
# how_many = 10 : show first ten results

most_skipped(streaming_history, track_or_artist = "track", for_view = T, how_many = 10)
```

## measure_how_long_listened

This function calculates how long the user listened to spotify in given time period. It can give the result as the duration or as the percentage of the specified time period. 

### Example

```{r}
# streaming_history = streaming_history : part of streaming history of one user, made with complete_streaming_history
# start_date = "2018-01-01" : filter the data table to only include results later than "2018-01-01"
# end_date = "2019-01-01" : filter the data table to only include results earlier than "2019-01-01"
# as_percentage = TRUE : show the result as a percentage of the input time period from strat_date to end_date
# for_view = TRUE : show results in format good for reading but not the best for further computations 
# 
measure_how_long_listened(streaming_history, start_date = "2018-01-01", end_date = "2019-01-01", as_percentage = TRUE, for_viewing = TRUE)
```


## count_skipped

This function counts how many tracks were skipped (their play time was less than 10s) in given time period. Either as a number or as a percentage of all tracks played in given time period.

### Example

```{r}
# streaming_history = streaming_history : part of streaming history of one user, made with complete_streaming_history
# start_date = "2018-01-01" : filter the data table to only include results later than "2018-01-01"
# end_date = "2019-01-01" : filter the data table to only include results earlier than "2019-01-01"
# as_percentage = TRUE : show the result as a percentage of all tracks played from strat_date to end_date


count_skipped(streaming_history, start_date = "2018-01-01", end_date = "2019-01-01", as_percentage = TRUE)
```

## longest_session 

This function finds the longest listening session (see `continuous_listening`) and makes the data table with it.

### Example

```{r}
# streaming_history = streaming_history : streaming_history : part of streaming history of one user, made with complete_streaming_history
# mins = 5 : the breaks must be less than 5 minutes
long_session <- longest_session(streaming_history, mins = 5)

#see the data
head(long_session)
```

## make_session_stats

This function calculates and returns the stats of the single listening session.

### Example

```{r}
# session_dt = long_session : see the stats of the longest session found earlier with longest_session function 
make_session_stats(long_session)
```

## session_length

This function shows about length of every session. Excludes skipped tracks.

### Example

```{r}
# streaming_history = streaming_history : streaming_history : part of streaming history of one user, made with complete_streaming_history
# mins = 5 : the breaks must be less than 5 minutes
# see the data
head(sessions_length(streaming_history, mins = 5))
```

## session_intervals

This function makes and returns the table with sessions divided into listening intervals.  
Sessions length are divided into categories: 

* <10min  
* 10-30min  
* 0.5-1h  
* 1-2h  
* 2-4h  
* 4-6h  
* 6h+  

```{r}
# streaming_history = streaming_history : streaming_history : part of streaming history of one user, made with complete_streaming_history
# mins = 5 : the breaks must be less than 5 minutes

# see the table

head(sessions_intervals(streaming_history, mins = 5))
```

# Plots

## plot_searches

This function plots number of search queries performed by user on different days. As additional information it can show either from which platform or in which country did the user perform the search.

### Example

```{r}
# search_queries = search_queries : part of search queries of one user, made with search_qeuries_dt
# additional = "platform" : show which platform the user used
plot_searches(search_queries, additional = "platform")
```


## plot_in_playlists_count

This function plots the number of tracks or the time played for tracks from users playlist and other tracks. Can be shown as a percentage.

## Example
```{r}
# str_his_with_playlist_wide = playlists_wide : use the wide format of the data table with streaming history and playlists
# as_percentage = FALSE : show as a value and not as a percentage
# time_or_count = "time" : count the sum of playtimes
# time_unit = "hours" : set the unit of time shown to hours
plot_in_playlists_count(playlists_wide, as_percentage = FALSE, time_or_count = "time", time_unit = "hours")
```


## plot_playlist_popularity

This function plots the popularity of  the playlists - either counting the number of time  tracks from it were played or the time they were played.

### Example 

```{r}
# str_his_with_playlist_long = playlists_long : use the long format of the data table with streaming history and playlists
# time_or_count = "time" : count the sum of playtimes
# time_unit = "hours" : set the unit of time shown to hours

plot_playlist_popularity(playlist_long, time_or_count = "time", time_unit = "hours")
```

## plot_track_count

This function plots how many tracks were played overtime, split by some period of time.

# Example 

```{r}
# streaming_history =  streaming_history : part of streaming history of one user, made with complete_streaming_history
# only_skipped = FALSE : show skipped and played tracks
# by = "day" : split the dateline by day
plot_track_count(streaming_history, only_skipped = FALSE, by = "day")
```

## plot_track_count_by_period

This function plots either daily or hourly habits of listening of the user.

# Example

```{r}
# streaming_history =  streaming_history : part of streaming history of one user, made with complete_streaming_history
# include_skipped = FALSE : don't show skipped tracks
# period = "hour" : show hourly habbits
# as_percentage = FALSE : don't as the percentage
# by_weekday = TRUE : split the plot by weekdays
plot_track_count_by_period(streaming_history, period = "hour", as_percentage = FALSE, include_skipped = FALSE, by_weekday = TRUE)
```

##sessions_visualize
This function plots the number of sessions for different session lengths.Sessions length are divided into categories: 

* <10min  
* 10-30min  
* 0.5-1h  
* 1-2h  
* 2-4h  
* 4-6h  
* 6h+ 

```{r}
# streaming_history =  streaming_history : part of streaming history of one user, made with complete_streaming_history
# mins = 5 : the breaks must be less than 5 minutes
# as_percentage = TRUE : show as a percentage of the number of all sessions
sessions_visualize(streaming_history, mins = 5, as_percentage = TRUE)

```

