
---
title: "Basic of SpotifyViz"
author: "."
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic of SpotifyViz}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



This vignette introduces to basic analyze data from spotify using spoitfyviz package.

To begin, we'll load spotifyviz.

```{r, warning=F}
library(spotifyviz)
```

## Loading data


To analyze data from spotify first we have to load the data. We're gonna use `make_streaming_history_dt()` to load streaming history, `prepare_playlists()` to load playlists and `make_search_queries_dt()` for search queries. In addition, if we want to have table with streaming history suited for rest of the functions, we need to call `prepare_streming_history()`.

```{r, warning=F}

# str_his = make_streaming_history_dt("path")
# str_his_dt = prepare_streaming_history(str_his)
# 
# search_queries = make_search_queries_dt("path")
# 
# playlists = prepare_playlists("path")
```

Now we can see how do they look like.

```{r}
# head(str_his_dt)
# 
# head(search_queries)
# 
# head(playlists)
```

## Operations on data

Now, since we have our data in appropriate, we can start analyzing.

```{r}
#make_summary_dt(str_his_dt, start_date = "2018-01-01", end_date = "2019-01-01", as_percentage = T)
```


```{r}
#most_played(str_his_dt, track_or_artist = "artist", show_skipped = F, how_many = 10)
```

```{r}
# most_skipped(str_his_dt, track_or_artist = "track", for_view = T, how_many = 10)
```


```{r}
#how_long_listened(str_his_dt, start_date = "2018-01-01", end_date = "2019-01-01", as_percentage = T, for_viewing = T)
```

```{r}
#how_many_skipped(str_his_dt, start_date = "2018-01-01", end_date = "2019-01-01", as_percentage = T)
```


```{r}
#plot_searches(search_queries, additional = "platform")
```


```{r}
# playlists_wide = str_his_with_playlists_wide(playlists, str_his_dt)
# 
# plot_in_playlists_count(playlists_wide, as_percentage = T, time_or_count = "time", time_unit = "hours")
```


```{r}
# playlist_long = str_his_with_playlists_long(playlists, str_his_dt)
# 
# plot_playlist_popularity(playlist_long, time_or_count = "time", time_unit = "hours")
```


```{r}
# plot_track_count(str_his_dt, only_skipped = F, by = "week")
```

```{r}
# plot_track_count_by_period(str_his_dt, period = "hour", as_percentage = T, include_skipped = T, by_weekday = T)
```


```{r}
# sessions = continuous_listening(str_his_dt, mins = 5)
# head(sessions)


```

```{r}
# long_session = longest_session(str_his_dt, mins = 5)

# head(long_session)
```

```{r}
# make_session_stats(long_session)
```


```{r}

# head(sessions_length(str_his_dt, mins = 5))

# head(sessions_intervals(str_his_dt, mins = 5))
```


```{r}
# sessions_visualize(str_his_dt, mins = 5, as_percentage = T)

```

